News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-22 rupnik2015.tex
everything except evaluation

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

@@

\title{Event Tracking Across Languages - 
Cross-Lngual Document Similarity and Event Tracking \todo{Better title...}}

@@

\end{abstract}
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item Cross lingual document similarity (model)
%\item  Event Detection
%\item  Combing the approaches
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would liek to understand events ans sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%\end{itemize}

@@

\section{System Description}

@@

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log( \frac{N}{DF_k} )$, where $DF_k$ is the number of documents in the corpus which contain term $k$.  A document $TFIDF$ vector is its original vector multiplied element-wise by the weights.

The $TFIDF$ weighted vector space model document representation corresponds to a map $\phi : \text{text} \rightarrow \RR^n$ defined by:
$$\phi(d)_k = {TF}_k(d) \log( \frac{N}{{DF}_k}).$$

@@

\subsection{Datasets}

\section{Multilingual Latent Factor Models}\label{sec:models}

\subsection{Multilingual dataset}

@@

\subsection{Algorithms}
\label{sec:algorithms}

We will present three approaches: a simple approach based on $k$-means clustering, a standard approach based on singular value decomposition and a new method that is applicable in more general
settings. All approaches are based on linear maps and we do not focus on alternatives, such as machine translation and probabilistic models. Applicability of machine translation relies
on the availability of training resources (strongly aligned translation corpora), which are not available for many world languages, but there exist comparable corpora. Probabilistic
methods \todo{provide reasons for not using them: e.g. difficulty with parameter estimation}.

\subsubsection{$k$-means}

The $k$-means algorithm is perhaps the most well-known and used clustering algorithm. In order to apply the algorithm we first merge all the term-document matrices into a single matrix $X$ by stacking the individual term-document matrices:
$$X := \begin{bmatrix}X_1^T ,X_2^T, \cdots, X_m^T \end{bmatrix}^T,$$
such that the columns respect the alignment of the documents (here MATLAB notation for concatenating matrices is used). Therefore, each document  is represented by a long vector indexed by the terms in all languages.

@@

\subsection{Canonical Correlation Analysis}
 For completeness we now introduce a statistical technique to analyze data from two sources - after which we will describe our main contribution to computing cross-lingual similarities.

@@

In this section we describe a method that is based on linear embeddings as well, but is more applicable to a large number of languages. The main difficulty with applying the
LSI and $k$-means approaches lies in the fact, that when one considers a large number of languages, for example, the top 100 Wikipedia languages (ranked by number of articles), the set of completely
aligned documents can be small or empty. Even if only two languages are considered, the set of aligned documents can be small (for example, Piedmontese and Hindi
Wikipedias had no interlanguage links). In Wikipedia we observed that even though the training resources are scarce between certain language pairs, often there exists indirect training data
through a hub language.

A \emph{hub language} is a language with a high proportion of non-empty documents in $D = \left\{d_1,..., d_{\ell}\right\}$, which in case of Wikipedia is English. We use the following notation to define subsets of the multilingual comparable corpus: let $a(i,j)$ denote the index set of all multilingual documents with non-missing data for the $i$-th and $j$-th language:  $a(i,j) = \left\{k | d_k = (u_1,...,u_m), u_i \neq \emptyset, u_j \neq \emptyset \right\},$ and let $a(i)$ denote the index set of all multilingual documents with non missing data for the $i$-th language.

@@

\subsection{Scaling}

100 languages!
Describe the newsfeed volume, number of events and the number of similarities that need to be computed. Compare the memory requirements of our current implementation to the rest.
Comment on the caching gains in our implementation.

\subsection{Evaluation}

\subsubsection{Link prediction}

\subsubsection{Hub language}

{\oldText{Taken from XLite paper}
To investigate the empirical performance of our algorithm we will test it on a large-scale, real-world multilingual dataset that we extracted from Wikipedia  by  using so called 'inter-language' links as an alignment. We select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing we remove the documents that contain less than 20 different words (stubs) and remove words occur in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors. These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too infrequent as well as the top 100 most frequent words in the vocabularies. Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.

@@

\section{Cross-lingual Event Linking}

One of the use cases where article linking information can be applied is also in an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles. The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''. Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015. Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

The clustering of articles is done for each language separately -- English articles are, for example, clustered separately from Spanish articles. When events are identified, they will therefore be identified from clusters in each language separately. As a result, the same event will be represented in Event Registry with two or more event IDs, and each event will only contain coverage in a single language. To avoid this problem we want to exploit the available information about the similarity of the articles in a cluster with articles in other languages. Using this information we hope to be able to connect clusters in different languages and represent them using a single event ID.

Formally, the problem of identifying the clusters describing the same event can be formalized as follows. Each article $a_i$ is written in a language $l$, where $l \in L = \{l_1,l_2,...,l_k\}$. For each language $l$, a set of clusters $C_l$ is generated and maintained that only contains articles in language $l$. For any cluster $c_i \in C_{l_a}$, the task is then to identify other clusters $c_j \in C_{l_b}$ (where $l_a \neq l_b$) that are describing the same event as $c_i$. We will refer to cluster pairs $c_i$ and $c_j$ that are describing the same event as equivalent clusters. An additional requirement of the task is that identification of the equivalent clusters should not be done exhaustively -- for a cluster $c_i$ we wish to avoid testing all other clusters $c_j$. Performing exhaustive comparison for each cluster $c_i$ would result in $O(N^2)$ tests, where $N$ is the number of all clusters, which is not feasible when the number of clusters is in tens of thousands.

In order to identify clusters $c_j$ that are equivalent to cluster $c_i$ we have developed a procedure that consists of two main tasks. The first task is to efficiently identify a limited set of clusters that are potential candidates for being equivalent to $c_i$. The second task is then about performing a detailed analysis of the candidates and finding the ones that are equivalent to $c_i$.

The details of the first task are described in Algorithm~\ref{cluster_merge_algo1}. The algorithms starts by individually inspecting each article $a_i$ in the cluster $c_i$. Using the CCA it then identifies 10 most similar articles to $a_i$ in each language $l \in L$. For each identified similar article $a_j$ we then identify cluster $c_j$ to which the article belongs to and add the cluster to the set of candidates that should be more closely inspected. The assumption made by this algorithm is that if two articles in different languages are describing the same event, the CCA similarity between the documents will be higher than when comparing two articles about different events. Since the CCA similarity is only an estimate, the assumption does not always hold. Due to data redundancy, however, we do not consider this to be an issue. Since we are testing each article in the cluster and compute 10 most similar articles for each article, we are very likely to identify among the candidates also the clusters that are equivalent to cluster $c_i$.

The second part of the procedure is described in the Algorithm~\ref{cluster_merge_algo2}. The goal of this task is to determine which (if any) of the candidate clusters are equivalent to $c_i$. To solve the task we represent it as a supervised learning problem. For each candidate cluster $c_j \in C'$ we compute a vector of learning features that should be indicative of whether the $c_i$ and $c_j$ are equivalent or not. The details about these features will be described in the next section. Once the features are computed, we classify the example with a classification model that was trained using a training dataset of hand-labeled examples. If the model predicts that tested clusters are equivalent, the $c_j$ is added to the set $C'$.

In order to train a classification model for predicting whether two clusters are equivalent we first need labeled training data. To obtain the data we used two human annotators. They were provided with a user interface that was displaying information about the articles in two clusters and their task was to determine if the clusters are equivalent or not. To obtain a pair of clusters $(c_i, c_j)$ to annotate, we first randomly chose a cluster $c_i$, used the Algorithm~\ref{cluster_merge_algo1} to compute a set of potentially equivalent clusters $C$ and randomly chose a cluster $c_j \in C$.

The dataset provided by the annotators contains 808 examples, of which 402 are equivalent clusters pairs and 406 are not. Clusters in each learning example are either in English, Spanish or German language. Although Event Registry imports articles also in other languages, we restricted our experiments to these three languages since they have the highest volume of news articles which makes the training problem similarly hard.

To build a classification model we first needed to extract from the labeled examples informative features. Based on the availability of the data, we have extracted the following groups of features:

\begin{itemize}
\item {\bf Cross-lingual article linking features}. Cross-lingual article linking information was already used to obtain candidates for the equivalent clusters. Additionally, it can also be used to provide valuable learning features. There are two main features that we extract from article links between clusters -- \texttt{linkCount} and \texttt{avgSimScore}. \texttt{linkCount} is the number of times an article $a_i \in c_i$ has as one of the 10 most similar articles an article $a_j \in c_j$. In other words, it is the number of times an article from $c_i$ has a very similar article, which is in $c_j$. Beside the number of links between the two clusters we can also take into account the similarity scores of the links. The \texttt{avgSimScore} is the feature that represents the average similarity score of the links between the two clusters.

\item {\bf Concept related features}. As mentioned before, articles that are imported into Event Registry are first semantically annotated by linking mentioned entities and keywords to the corresponding Wikipedia pages. Whenever Barack Obama is, for example, mentioned in the article, the article is annotated with a link to his Wikipedia page (\emph{http://en.wikipedia.org/wiki/Barack\_Obama}). In the same way all mentions of people, locations, organizations and even ordinary keywords (e.g. bank, tax, ebola, plane, company) are annotated. Although the Spanish article about Obama will be annotated with his Spanish version of the Wikipedia page, we can almost in all cases normalize the Wikipedia links to their English versions. This can be done since Wikipedia itself provides information regarding which pages in different languages represent the same concept/entity. Using this approach, the word ``avi\'on'' in a Spanish article will be annotated with the same concept as the word ``plane'' in an English article. Although articles are in different languages, the annotations can therefore provide a shared vocabulary that can be used to compare articles/clusters. By analyzing all articles in clusters $c_i$ and $c_j$ we can identify the most relevant entities and keywords for each cluster. Additionally we can also assign weights to the concepts based on how frequently they occur in the articles in the cluster. From the list of relevant concepts and corresponding weights we can now define a number of features. \texttt{entityCosSim} represents the cosine similarity on the vector of entities mentioned in both clusters. Similarly, \texttt{keywordCosSim} represents the cosine similarity on the keywords only. In case the weights are not relevant, we have also added two features (\texttt{entityJaccardSim} and \texttt{keywordJaccardSim}), where Jaccard similarity is used instead of the cosine similarity.

\item {\bf Miscellaneous features}. This group contains three miscellaneous features that seem discriminative but are unrelated to the previous two groups. The first feature is related to the event location. By analyzing the locations mentioned in the articles in a cluster we are able to estimate where the event occurred. The \texttt{hasSameLocation} feature is a boolean variable that is true when the location of the event in both clusters is the same. Another important feature is related to the time when the articles in the clusters were published. For articles in each cluster we first compute the average date and time when the articles were published. The \texttt{timeDiff} is then defined as the absolute difference in hours between the two computed dates. The last feature in this group is computed by analyzing the mentions of dates in the articles. Using an extensive set of regular expressions we are able to detect mentions of dates in different forms. To compute the feature \texttt{sharedDates} we start by identifying the list of dates mentioned in articles in each cluster separately. The value of \texttt{sharedDates} is then determined as the ratio of all dates that are detected in both clusters.

Using the described features and the user-provided label we can train a classification model $M$ that can be used in Algorithm~\ref{cluster_merge_algo2} to classify a new cluster pair as equivalent or not. The classification algorithm that we used to train a model was a linear SVM method.

\subsection{Evaluation}

@@

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{comment}

@@

\title{News Across Languages -
Cross-Lingual Document Similarity and Event Tracking}

@@

\end{abstract}
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item Cross lingual document similarity (model)
%\item  Event Detection
%\item  Combing the approaches
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would liek to understand events ans sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%%\end{itemize}
%
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item System overview
%\item Cross lingual document similarity (algorithms)
%\item Event linking
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would like to understand global news as sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%\end{itemize}
%

@@

\section{System Overview}

Event registry $\Rightarrow$ Monolingual events $\Rightarrow$ features (semantic + CCA) $\Rightarrow$ linked events (crosslingual)

@@

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$.  A document $TFIDF$ vector is its original vector multiplied element-wise by the weights.

The $TFIDF$ weighted vector space model document representation corresponds to a map $\phi : \text{text} \rightarrow \RR^n$ defined by:
$$\phi(d)_k = {TF}_k(d) \log\left( \frac{N}{{DF}_k}\right).$$

@@

\section{Multilingual Latent Factor Models}\label{sec:models}
In this section, we describe several approaches to computing multilingual simliarities described in the previous section. We present three approaches
\begin{enumerate}
\item  a simple approach based on $k$-means clustering
\item  a standard approach based on singular value decomposition 
\item a new method that is applicable in more general settings.
\end{enumerate}
 We concentrate on approaches that are based on linear maps rather than alternatives, such as machine translation and probabilistic models. Applicability of machine translation relies
on the availability of training resources (strongly aligned translation corpora), which are not available for many world languages, but there exist comparable corpora. Probabilistic
methods \todo{provide reasons for not using them: e.g. difficulty with parameter estimation, maybe just say future work?}.

For completeness we will also present the method of Canonical Correlation Analysis, since it is related to our proposed method.

\subsection{Notation}

@@

\subsection{$k$-means}

The $k$-means algorithm is perhaps the most well-known and widely-used clustering algorithm. In order to apply the algorithm, we first merge all the term-document matrices into a single matrix $X$ by stacking the individual term-document matrices:
$$X := \begin{bmatrix}X_1^T ,X_2^T, \cdots, X_m^T \end{bmatrix}^T,$$
such that the columns respect the alignment of the documents (here MATLAB notation for concatenating matrices is used). Therefore, each document  is represented by a long vector indexed by the terms in all languages.

@@

\subsection{Canonical Correlation Analysis}
 The final approach is a statistical technique to analyze data from two sources - after which we will describe our main contribution to computing cross-lingual similarities.

@@

In this section, we describe an extension to CCA, but is more applicable to a large number of languages. The main difficulty with applying the
LSI and $k$-means approaches lies in the fact, that when one considers a large number of languages, for example, the top 100 Wikipedia languages (ranked by number of articles), the set of completely
aligned documents is often small or empty. Even if only two languages are considered, the set of aligned documents can be small (for example, Piedmontese and Hindi
Wikipedias had no interlanguage links). In Wikipedia, we observed that even though the training resources are scarce between certain language pairs, there often exists indirect training data
through what we refer to as a hub language.

A \emph{hub language} is a language with a high proportion of non-empty documents in $D = \left\{d_1,..., d_{\ell}\right\}$. The prototypical  exampling  in the case of Wikipedia is English. We use the following notation to define subsets of the multilingual comparable corpus: let $a(i,j)$ denote the index set of all multilingual documents with non-missing data for the $i$-th and $j$-th language:  $a(i,j) = \left\{k | d_k = (u_1,...,u_m), u_i \neq \emptyset, u_j \neq \emptyset \right\},$ and let $a(i)$ denote the index set of all multilingual documents with non missing data for the $i$-th language.

@@

(figure?)

\section{Cross-lingual Event Linking}
The main application we test the above simliarity on  is cross-lingual event linking. In online media streams - particularly new articles - there is often duplication of reporting, different viewpoints or opinions, all centering around a single event.  Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015. (Sports example)

Informally, events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry\cite{}. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to find suitable matchings between clusters with articles in different languages.

\begin{comment}
One of the use cases where article linking information can be applied is also in an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles. The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''.

The clustering of articles is done for each language separately -- English articles are, for example, clustered separately from Spanish articles. When events are identified, they will therefore be identified from clusters in each language separately. As a result, the same event will be represented in Event Registry with two or more event IDs, and each event will only contain coverage in a single language. To avoid this problem we want to exploit the available information about the similarity of the articles in a cluster with articles in other languages. Using this information we hope to be able to connect clusters in different languages and represent them using a single event ID.
\end{comment}

The problem in the second phase of the pipeline is to identify the clusters (given?) describing the same events. Each article $a_i$ is written in a language $l$, where $l \in L = \{l_1,l_2,...,l_k\}$. For each language $l$, we generate and maintain a set of language-specific clusters $C_l$. That is, all articles are in language $l$. For any cluster $c_i \in C_{l_a}$, we would like to identify clusters $c_j \in C_{l_b}$ (where $l_a \neq l_b$) that describe the same event as $c_i$. We define \emph{equivalent clusters} as  pairs $(c_i,c_j)$ such that both clusters in the pair describe the same event in languages $C_{l_a}$ and $C_{l_b}$ respectively.

We impose an additional requirement.  Identification of the equivalent clusters cannot be done exhaustively -- for a cluster $c_i$ we wish to avoid testing all other clusters for all the other languages. Performing exhaustive comparison for each cluster $c_i$ would result in $O(N^2)$ tests, where $N$ is the number of all clusters (over all languages), which is not feasible for when the number of clusters is on the order of tens of thousands.

\subsection{Related work}

TODO

In order to identify clusters $c_j$ that are equivalent to cluster $c_i$, we have developed a two-stage algoirhtm. First, for a cluster $c_i$, we efficiently identify a small set of candidate clusters and then  find the clusters from among the candidates which are equivalent to $c_i$.

The details of the first step are described in Algorithm~\ref{cluster_merge_algo1}. The algorithm begins by individually inspecting each article $a_i$ in the cluster $c_i$. Using CCA  (see Section~\ref{}),  it then identifies the 10 most similar articles to $a_i$ in each language $l \in L$. For each  similar article $a_j$, we identify its corresponding  cluster $c_j$ and add it to the set of candidates. \todo{we should say a few words why this is more efficient - since now we are comparing each article to all other articles - while we said it is not feadible for clusters}. While the CCA simliarity is approximate, our  assumption  is that articles in different languages describing the same event, will generally have a higher similarity than articles about different events. While this assumption does not always hold, redundancy in the data mitigates these false positives. Since we compute the 10 most similar articles for each article in $c_i$, we are likely to identify all the relevant candidates for cluster $c_i$.

The second stage of the algorithm (described in the Algorithm~\ref{cluster_merge_algo2}.) determines which (if any) of the candidate clusters are equivalent to $c_i$. We approach this as a supervised learning problem. For each candidate cluster $c_j \in C'$, we compute a vector of learning features that should be indicative of whether the $c_i$ and $c_j$ are equivalent or not. The details of our chosen features will be described in the following Section. Once these features are computed, we apply a classification model which was trained using a  dataset  of hand-labeled examples. If the model predicts that tested clusters are equivalent, the $c_j$ is added to the set $C'$.

In order to train a classification model for predicting whether two clusters are equivalent, we must first have labeled data. To obtain the data, we used two human annotators. The annotators were provided with an interface displaying information about the articles in two clusters and their task was to determine if the clusters were equivalent or not. To obtain a pair of clusters $(c_i, c_j)$ to annotate, we first randomly chose a cluster $c_i$, used the Algorithm~\ref{cluster_merge_algo1} to compute a set of potentially equivalent clusters $C$ and randomly chose a cluster $c_j \in C$.

The dataset provided by the annotators contains 808 examples, of which 402 are equivalent clusters pairs and 406 are not. Clusters in each learning example are either in English, Spanish or German. Although Event Registry imports articles in other languages as well, we restricted our experiments to these three languages, since they have the highest volume of articles which makes the training problem similarly hard \todo{doesnt this make it easier?}.

To build a classification model we first extract  informative features from the labeled examples. Based on the availability of the data, we extracted the following groups of features:

\begin{itemize}
\item {\bf Cross-lingual article linking features}. Cross-lingual article linking information was already used to obtain candidates for the equivalent clusters. Additionally, it can also be used to provide valuable  features for learning. There are two main features that we extract from article links between clusters -- \texttt{linkCount} and \texttt{avgSimScore}. \texttt{linkCount} is the number of times an article $a_i \in c_i$ has as one of the 10 most similar articles an article $a_j \in c_j$. In other words, it is the number of times an article from $c_i$ has a very similar article, which is in $c_j$. Beside the number of links between the two clusters we can also take into account the similarity scores of the links. The \texttt{avgSimScore} is the feature that represents the average similarity score of the links between the two clusters.

\item {\bf Concept related features}. As mentioned before, articles that are imported into Event Registry are first semantically annotated by linking mentioned entities and keywords to the corresponding Wikipedia pages. Whenever Barack Obama is, for example, mentioned in the article, the article is annotated with a link to his Wikipedia page (\emph{http://en.wikipedia.org/wiki/Barack\_Obama}). In the same way all mentions of people, locations, organizations and even ordinary keywords (e.g. bank, tax, ebola, plane, company) are annotated. Although the Spanish article about Obama will be annotated with his Spanish version of the Wikipedia page, in many cases we can link the Wikipedia pages to their English versions. This can be done since Wikipedia itself provides information regarding which pages in different languages represent the same concept/entity. Using this approach, the word ``avi\'on'' in a Spanish article will be annotated with the same concept as the word ``plane'' in an English article. Although the articles are in different languages, the annotations can therefore provide a shared vocabulary that can be used to compare articles/clusters. By analyzing all the articles in clusters $c_i$ and $c_j$, we can identify the most relevant entities and keywords for each cluster. Additionally, we can also assign weights to the concepts based on how frequently they occur in the articles in the cluster. From the list of relevant concepts and corresponding weights, we consider the following features: \texttt{entityCosSim} which represents the cosine similarity on the vector of entities mentioned in both clusters; \texttt{keywordCosSim} which represents the cosine similarity on the keywords only; for cases where the weights are not relevant, we have also added two additional features - (\texttt{entityJaccardSim} and \texttt{keywordJaccardSim}), where Jaccard similarity is used instead of the cosine similarity.

\item {\bf Miscellaneous features}. This group contains three miscellaneous features that seem discriminative but are unrelated to the previous two groups. The first feature is related to the event location. By considering the locations mentioned in the articles in a cluster, we are able to estimate where an event occurred. The \texttt{hasSameLocation} feature is a boolean variable that is true when the location of the event in both clusters is the same. Another important feature is related to the time when the articles in the clusters were published. For each cluster, we first compute the average publication time and date of the articles. The \texttt{timeDiff} is then defined as the absolute difference in hours between the two computed dates. The final feature we consider is computed by analyzing the mentions of dates in the articles. Using an extensive set of regular expressions, we are able to detect mentions of dates in different forms. To compute the feature \texttt{sharedDates} we start by identifying the list of dates mentioned in articles in each cluster separately. The value of \texttt{sharedDates} is then determined as the ratio of all dates that are detected in both clusters.\todo{this should be made more explicit/}

Using the  features described above and the user-provided labels, we train a classification model $M$ which is used in Algorithm~\ref{cluster_merge_algo2} to classify a new cluster pair as equivalent or not. The classification algorithm that we used to train a model was a linear SVM method.

\section{Evaluation}

We describe the datasets involved in the experiments and two evaluations: comparing the crosslingual approaches on an information retrieval task, and evaluation of cross-lingual event linking.

\subsection{Datasets}
\emph{Andrej started shuffling it, will do it till Monday :) }

\subsection{Wikipedia Information Retrieval}

Mate retrieval results

\subsection{Experiments With Missing Alignment Data}
\emph{Andrej started working on this, done by Monday}
{\oldText{Taken from XLite paper}
To investigate the empirical performance of our algorithm we will test it on a large-scale, real-world multilingual dataset that we extracted from Wikipedia  by  using so called 'inter-language' links as an alignment. We select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing we remove the documents that contain less than 20 different words (stubs) and remove words occur in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors. These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too infrequent as well as the top 100 most frequent words in the vocabularies. Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.

@@

\subsection{Evaluation Of Cross-Lingual Event Linking}

@@

\subsection{Scaling}
\emph{Andrej asks who is the rest? I think that this section should be short.}
100 languages!
%Newsfeed volume
Describe the newsfeed volume, number of events and the number of similarities that need to be computed. Compare the memory requirements of our current implementation to the rest.
Comment on the caching gains in our implementation.
